Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-05 submission.tex
endringer i sec:prev

\section{Previous Work}
\label{sec:prev}
There has been little previous work dedicated specifically to
extrinsic evaluation of the effects of PoS tag sets on downstream
applications.  There has, however, been some work that evaluates the
effects of PoS tag set granularity on PoS tagging. This includes
investigation of the effects of PoS tag sets on tagging of Swedish
\cite{Meg:01,Meg:02} and English \cite{Mac:05}.

@@

\newcite{Meg:02} trained and evaluated a range of PoS taggers on the
Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, they
investigated the effects of tag set size on tagging by mapping the original tag
set into smaller subsets designed for parsing. They argue that a tag set with
complete morphological tags may not be necessary for all NLP applications, for
instance syntactic parsing. They found that the smallest tag set comprising 26
tags yields the lowest tagger error rate.  However, for some of the taggers,
augmenting the tag set with more linguistically informative tags may actually
lead to a drop in error rate. They argue that this shows that the size of the
tag set as well as the type of information in the tags are crucial factors for
tagger performance.  However/unfortunately, they do not report results of
parsing with the various PoS tag sets.

Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
performance in English, specifically the Wall Street Journal portion of Penn
Treebank \cite{Mar:San:Mar:93}. They mapped the original tag set of Penn
Treebank comprising 45 tags to more fine-grained tag sets using linguistic
insight to investigate whether additional linguistic information included in
finer-grained tags could assist the tagger. Experimenting with both lexically
and syntactically conditioned modifications such as distinguishing between
count nouns and noncount nouns and between transitive and intransitive verbs,
they found that more fine-grained tag sets rarely led to improvements in tagger
accuracy; the most successful modification yielded an improvement in tagger
accuracy of 0.05 percentage points.  They did not find any statistically
significant improvements using linguistically informed distinctions, arguing
that their results do not support the hypothesis that it is possible to achieve
significant performance improvements in PoS tagging by utilizing a
finer-grained tag set.

Transition/overgang her ...

>>>>

\section{Previous Work}
\label{sec:prev}
There has been little previous work dedicated specifically to
extrinsic evaluation of the effects of PoS tag sets on downstream
applications.  There has, however, been some work that evaluates the
effects of tag set granularity on PoS tagging itself. This includes
investigation of the effects of tag sets on PoS tagging for Swedish
\cite{Meg:01,Meg:02} and English \cite{Mac:05}.

@@

\newcite{Meg:02} trained and evaluated a range of PoS taggers on the
Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, the effects 
of tag set size on tagging was investigated by mapping the original tag
set into smaller subsets designed for parsing. \newcite{Meg:02} argues that a tag set with
complete morphological tags may not be necessary for all NLP applications, for
instance syntactic parsing. The study found that the smallest tag set comprising 26
tags yields the lowest tagger error rate.  However, for some of the taggers,
augmenting the tag set with more linguistically informative tags may actually
lead to a drop in error rate. \newcite{Meg:02} argues that this shows that the size of the
tag set as well as the type of information in the tags are crucial factors for
tagger performance.  Unfortunately, results for parsing with the various PoS
tag sets are not reported.

Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
performance in English, specifically the Wall Street Journal portion of the Penn
Treebank (PTB) \cite{Mar:San:Mar:93}. Based on linguistic
considerations \newcite{Mac:05} mapped the 45 tags of the original PTB tag set
 to more fine-grained tag sets to investigate whether additional linguistic information 
could assist the tagger. Experimenting with both lexically
and syntactically conditioned modifications such as distinguishing between
count nouns and noncount nouns and between transitive and intransitive verbs,
they found that more fine-grained tag sets rarely led to improvements in tagger
accuracy; the most successful modification yielded an improvement in tagger
accuracy of 0.05 percentage points.  They did not find any statistically
significant improvements using linguistically informed distinctions, arguing
that their results do not support the hypothesis that it is possible to achieve
significant performance improvements in PoS tagging by utilizing a
finer-grained tag set.

% fixme:
Transition/overgang her ...
